Constructive Preference Elicitation by Setwise Max-margin Learning

### 2016-01-27 paper/ijcai16.tex
paolo:finished pass section 2

The assessment of preferences have received considerable attention in the literature, with early works confined to the operation research and management literature mostly \cite{}.
In the past decade a number of researchers have considered elicitation of preferences from an artificial intelligence perspective; within the machine learning field the term preference learning has been catching up.
While preferences are often assessed using some ad-hoc rules \cite{},
recently a number of researchers have proposed principled methods that elicit utility in an adaptive way \cite{chajewska2000,boutilier2002,Wang2003,boutilier2006,braziunas-mmr:uai07,guo2010real,viappiani2010optimal}; the aims is that of focusing on learning the ``important'' part of the utility, allowing to recommend good (or even optimal) decision with only a partial information about the utility function.

@@

Our approach is related to \cite{boutilier2006}, where minimax regret is used both
as a robust recommendation criterion and as a technique to drive elicitation in configuration problems.
The main limitation of the minimax regret approach is the lack of tolerance with respect to user inconsistency (the user might state that item x is preferred to y, even though the latter has higher utility).

% setwise max-margin formulation to deal with uncertainty in user utility
Indeed, learning a user utility function requires setting a preference
elicitation strategy and dealing with uncertain and possibly
inconsistent user feedback. 
Bayesian preference elicitation approaches
deal with both problems by building a probability distribution on
candidate functions (endowed with a response or error model to be used for inference) and asking queries maximizing informativeness
measures such as expected value of information (EVOI)~\cite{chajewska2000,viappiani2010optimal}.

@@

Consistently with previous work~\cite{guo2010real,viappiani2010optimal}, we model users by means
of linear utility functions: the user's preferences are represented by a weight
vector $\vw\in\bbR^m$, and the utility of a configuration $\vx$ is given by
$\langle \vw, \vx \rangle = \sum_{z=1}^m w_z x_z$. In the remainder of the paper we
require all weights to be {\em non-negative} and {\em bounded}: the per-attribute
weights $w_z$ must lie in a (constant but otherwise arbitrary) interval
$[w^\bot_z, w^\top_z]$, with $w^\bot_z \ge 0$. Both requirements are quite
natural, and enable the translation of our core optimization problem into a
mixed-integer linear problem (as done in Section~\ref{sec:formulation}).

@@

Users were simulated by drawing $20$ random utility vectors from each of four
different distributions. The first two mimic those used
in~\cite{guo2010real}\footnote{The normal distribution reported
in~\cite{guo2010real} differs slightly from the one implemented in the code of
their experimental setting. In this paper we use the latter.}:
(1) a uniform distribution over $[1, 100]$ for each individual weight, and (2)
a normal distribution with mean $25$ and covariance
$\frac{25}{3}$.  We further produced two novel
{\em sparse} sets of weight vectors by deleting $80\%$ of the entries of the
vectors sampled in the uniform and normal settings uniformly at random
\stefano{Paolo, is this correct?}. All methods were evaluated on the very same
weight vectors.

@@

>>>>

The assessment of preferences have received considerable attention in the literature, with early works confined to the operation research and management literature mostly \cite{}.
In the past decade a number of researchers have considered elicitation of preferences from an artificial intelligence perspective; within the machine learning field the term preference learning has been catching up.
While preferences are often assessed using some ad-hoc rules \cite{},
recently a number of researchers have proposed principled methods that elicit utility in an adaptive way \cite{chajewska2000,boutilier2002,Wang2003,boutilier2006,braziunas2007,guo2010real,viappiani2010optimal}; the aims is that of focusing on learning the ``important'' part of the utility, allowing to recommend good (or even optimal) decision with only a partial information about the utility function.

%braziunas-mmr:uai07,

@@

Preference elicitation in configuration problems has been previously tackled with regret-based elicitation \cite{boutilier2006,braziunas2007}, where minimax regret is used both
as a robust recommendation criterion and as a technique to drive elicitation in configuration problems.
The main limitation of the minimax regret approach is the lack of tolerance with respect to user inconsistency (the user might state that item x is preferred to y, even though the latter has higher utility).

% setwise max-margin formulation to deal with uncertainty in user utility
Indeed, learning a user utility function requires setting a preference
elicitation strategy and dealing with uncertain and possibly
inconsistent user feedback. 
Bayesian preference elicitation approaches
deal with both problems by building a probability distribution on
candidate functions (endowed with a response or error model to be used for inference) and asking queries maximizing informativeness
measures such as expected value of information (EVOI)~\cite{chajewska2000,guo2010real,viappiani2010optimal}.

@@

Consistently with previous work~\cite{guo2010real,viappiani2010optimal}, 
we model users by means of utility functions of a given structure; the user's preferences are represented by a latent weight
vector $\vw\in\bbR^m$.
We assume linear utility functions; the utility of a configuration $\vx$ is given by
$\langle \vw, \vx \rangle = \sum_{z=1}^m w_z x_z$
(see \cite{keeney1976} for theoretical conditions justifying additive utility functions). 
In the remainder of the paper we
require all weights to be {\em non-negative} and {\em bounded}: the per-attribute
weights $w_z$ must lie in a (constant but otherwise arbitrary) interval
$[w^\bot_z, w^\top_z]$, with $w^\bot_z \ge 0$. 
Both requirements are quite natural\footnote{Utility values are defined on an
interval scale, therefore the assumption of non negativity is not restrictive as it is always possible to scale the values appropriately (see for instance \cite{Torra2007} and \cite{keeney1976}).}, and enable the translation of our core optimization problem into a
mixed-integer linear problem (as done in Section~\ref{sec:formulation}).

@@

Users were simulated by drawing $20$ random utility vectors from each of four
different distributions. The first two mimic those used
in~\cite{guo2010real}\footnote{The normal distribution reported
in~\cite{guo2010real} differs slightly from the one implemented in the code of
their experimental setting. 
In our experiments we use the latter.}:
(1) a uniform distribution over $[1, 100]$ for each individual weight, and (2)
a normal distribution with mean $25$ and %covariance
standard deviation $\frac{25}{3}$ (each attribute is sample i.i.d).  
We further produced two novel
{\em sparse} sets of weight vectors by deleting $80\%$ of the entries of the
vectors sampled in the uniform and normal settings uniformly at random
\stefano{Paolo, is this correct?}. All methods were evaluated on the very same
weight vectors.

@@

\paolo{Maybe a plot of an experiment with increased noise in user reponse? (for example, $\beta=0.1$ instead of $1$.}
